---
title: "Diet Distribution Analysis"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(knitr)
library(kableExtra)
source("diet.R")
knitr::opts_chunk$set(echo = TRUE)
```
Can food trigger people's emotional and mental health? Absolutely. Many people are surprised that there is a close link between mental health disorders and nutritional deficiencies. The role of food is more prominent than people think! Bad eating and drinking habits can negatively impact people's thoughts and behaviors. Malnutrition can also lead to physical health problems, affecting people's feelings and way of thinking. Few people understand the close relationship between nutrition and mental health disorders such as depression and anxiety. Depression and anxiety are often regarded as mysterious diseases shrouded in shame. Depression is generally believed to be based on chemical imbalances or unstable emotions. However, surprisingly, nutrition often plays a significant role in the onset, duration, and intensity of depression. With that being said, the diet distribution analysis portion will compare and contrast the relationship between healthy diet patterns and mental health by observing obesity and a healthy diet.

<div class = "row">
<div class = "col-md-6">
```{r echo=FALSE}
top_5_obes %>%
  kbl(caption = "Top Five Countries with the Highest Obesity Percentage") %>%
  kable_styling(
    bootstrap_options = "striped",
    full_width = F, position = "left"
  )
```
</div>

##### The Recommended Simple Daily Diet Intake by The USDA Center for Nutrition Policy and Promotion
<div class = "col-md-6">
```{r echo=FALSE}
print(ideal_diet_plot)
```
</div>
</div>

Based on the data obtained from the Food and Agriculture Organization of the United Nations, the top five countries with the highest obesity rate are listed in the table (left). The plot on the right is a simple daily diet guideline recommended by USDA Center for Nutrition Policy and Promotion: 30% grains, 40% vegetables, 10% fruits, and 20% protein. Below, look at different categories of food intake in five obese countries: food supply quantity (kg), fat quantity, energy intake (kcal), and protein, to predict the relationship between healthy eating patterns, obesity rates, and mental health.


### Different Categories of Food Intake Quantities for the Top Five Obesity Countries
```{r, echo=FALSE, message=FALSE, out.width="50%"}
print(obes_kg_plot)
print(obes_fat_plot)
print(obes_kcal_plot)
print(obes_protein_plot)
```

The plots above, which display the four different food intake categories in the top five obesity countries, seem to be inconsistent with the recommended dietary guidelines. In terms of these five countries' total dietary intake quantity (kg), the sum of all four food categories (fruits, grains, protein, vegetables) is only about 40%. We can then infer that the remaining 60% of food intake is dairy products, oils, fat, sweets, and others. Therefore, we can predict that the 60% food intake is a significant contributor to their high obesity rate since more than half of what their people eat is considered "unhealthy." Nevertheless, according to the order of intake quantity of the four varieties of food, Kuwait had best follow this order, with the largest consumption of vegetables, followed by grains and protein, and finally fruits. The fat category shows that people in these five countries mainly get fat intake from protein. Although the total fat intake of these four categories in five countries is below 25%, it is still not good because the total intake quantity is only about 40%. As for energy (kcal) intake, this shows that they all get most of their energy (kcal) from grains. Saudi Arabia topped the list with 25% of its kcal intake from protein and followed by Kuwait with about 20%. Finally, people in these countries get their protein mainly from their protein intake, except Saudi Arabia, which gets its protein from grains. In conclusion, the study of four different food intake categories in the top five obesity countries shows a solid relationship between nutrition and health. Because the dietary patterns in these countries are far lower than the recommended dietary guidelines, the obesity rate is very high. In the case of such a high obesity rate, we can also think that the mental health status of these obese people is not necessarily the best because unhealthy eating habits can lead to depression and anxiety.
